---
title: "Mini-Project 2: Forecasting US Energy Consumption"
author: YOUR NAME HERE
format: 
  html: 
    embed-resources: true
editor: visual
---

You are asked to predict the monthly US residential energy consumption for the next 2 years for the 'Short-Term Energy Outlook' using data from the US Department of Energy.

```{r}
#| include: false
# load packages
library(tidyverse)
library(astsa)
```

## IMPORT

The US Department of Energy collects an impressive amount of data to support the energy industries. For this assignment use the monthly total energy consumed by the US residential sector. The webpage

<https://www.eia.gov/totalenergy/data/browser/index.php?tbl=T02.02>

contains monthly web tables, but an easier approach to getting the data into R uses the result of the 'Download' button to read the source data from the source CSV file.

```{r}
#| warning: false
energy <- read_csv("https://www.eia.gov/totalenergy/data/browser/csv.php?tbl=T02.02",
# names in first row, but the names aren't meaningful
                   col_names = c("MSN","Time","Res_Energy","label1","label2","label3"),
                   skip = 1) |>
  mutate(Res_Energy = parse_number(Res_Energy),
         Time = ym(Time))
```

## TIDY

```{r}
#| eval: false
"___"(energy)
```

Prompt: Why is `energy` tidy?

*Note:* Your answer requires a response to each of the 3 tidy data rules.


## TRANSFORM

Code should

-   filter to the data in rows where `MSN` is `TERCBUS` (Total Energy Consumed by the Residential Sector in trillion Btu),

-   filter to current data (after 1991),

-   remove the annual total rows (`Time` is missing / `NA`).

-   only include the columns `Time` and `Res_Energy` columns, and

```{r}
#| eval: false
energy <- energy |>
# filter to TERCBUS Total Energy Consumed by the Residential Sector
  filter("___") |>
# filter to "since the 90's"
  filter("___")) |>
# remove annual total rows (Time is NA)
  "___".omit() |>
# keep just the Time and Res_Energy columns
  select("___")

# take a look
tail(energy)
```


## VISUALIZE

Plot the time series for US residential energy consumption (in trillion Btu).

**Note:** You will need to set `eval: true` when you have created the `energy` dataset in the TRANSFORM section.

```{r}
#| eval: false
ggplot(energy, aes(x = "___", y = "___")) +
  geom_"___"() +
  labs(
    #title = "___",
    #x = "___", 
    #y = "___"
    )
```

## MODEL

Fit an ARIMA$(1, 1, 1) \times (1, 1, 1)_{12}$ model and report the parameter estimates and standard errors.

**Note:** You will need to set `eval: true` when you have created the `energy` dataset in the TRANSFORM section.

```{r}
#| eval: false
energy_out <- sarima(energy$Res_Energy, 1,1,1, 1,1,1, 12)

# table of estimates and se
energy_out$ttable
```

Forecast the US residential energy consumption for the next 2 years. Report a table of predictions and 95% prediction intervals.

**Note:** You will need to set `eval: true` when you have created the `energy` dataset in the TRANSFORM section.

```{r}
#| eval: false
#  "next two years" 
energy_forecast <- data.frame(Time = max(energy$Time) + months(1 : 24))
energy_yhat <- sarima.for(energy$Res_Energy , n.ahead = 24, 1,1,1, 1,1,1, 12)

# table of forecasts and uncertainty
energy_forecast$pred <- energy_yhat$pred
energy_forecast$L <- energy_yhat$pred - 2 * energy_yhat$se
energy_forecast$U <- energy_yhat$pred + 2 * energy_yhat$se

energy_forecast
```

## COMMUNICATE

**Prompt:** Does it appear there is a monthly pattern? If so, explain the possible business/economic 'cause' for this effect.
