---
title: "HW 2: Palmer Island Penguins"
author: YOUR NAME HERE
format: 
  html: 
    embed-resources: true
editor: visual
---

```{r}
#| label: load-packages
#| include: false
library(tidyverse)
library(palmerpenguins)
```

## 1. Palmer Penguins

Measurements for penguin species, island in Palmer Archipelago, size (flipper length, body mass, bill dimensions), and sex.

**Import:**

```{r}
penguins
```

**Visualize:**

Make a plot showing differences between penguins and their bills.

*Hint*: Look at the "Coding Out Loud" lecture for the graphic that answers this question.

*Note*: Replace the `"___"` and uncomment (remove the `#`) to see the effect

```{r}
ggplot(penguins,
       aes(x = bill_depth_mm,
           y = bill_length_mm,
           color = species)) +
  #geom_"___"() +
  labs(
    #"___" = "Bill depth and length",
       
    #"___" = "Dimensions for Adelie, Chinstrap, and Gentoo Penguins",
    #"___" = "Bill depth (mm)", "___" = "Bill length (mm)",
    #"___" = "Species",
    #"___" = "Source: Palmer Station LTER / palmerpenguins package"
    ) +
    scale_color_viridis_d() 
```

## 2. Chinstrap Penguins

Chinstrap penguins are one of the species in the Palmer Archipelago.

**Import:**

```{r}
chinstrap <- penguins |>
  filter(species == "Chinstrap") |>
  na.omit()
chinstrap
```

**Visualize flipper length:**

*Hint*: Since penguin flipper length is univariate data, to answer this question you will want to try a histogram, density plot, boxplot and your answer will be the one of them that you think best communicates the shape, center, variation, and anything unusual about the distribution.

*Note*: Change `eval` to `true` and replace the `"___"`

```{r}
#| eval: false
ggplot(chinstrap, aes(x = "___")) +
  geom_"___"("___") +
  labs(
    x = "___",
    y = "___",
    title = "___"
  )
```

**Communicate:**

Prompt: What should someone know about the Chinstrap penguin flipper length distribution? (shape, center, spread, anything unusual)

**Visualize distribution of year (data collection over time):**

*Note*: Change `eval` to `true` and replace the `"___"`

```{r}
#| eval: false
ggplot(chinstrap, aes(x = "___")) +
  geom_"___"("___") 
```

**Communicate:**

Prompt: What should someone know about how the Chinstrap penguin data was collected over time (different years)?

## Stretch Exercise:

Is the Chinstrap penguin flipper length distribution approximately normal?

(Insert R code cell and write appropriate code)
