---
title: "HW 3: Mortgage Rates"
author: YOUR NAME HERE
format: 
  html: 
    embed-resources: true
editor: visual
---

```{r}
#| label: load-packages
#| include: false
library(tidyverse)
library(scales)
```

## Mortgage Rates over time

Housing is often a barometer for how people feel about the economy. Interest rates for mortgages are also the lowest of the interest rates for other personal borrowing.

Freddie Mac, a publicly traded government-sponsored enterprise focused on providing an affordable supply of mortgages across the US, shares the weekly average 30-year mortgage rate. The average is over applications that week. 

**Import:**

```{r}
rate <- read_csv("https://fred.stlouisfed.org/graph/fredgraph.csv?id=MORTGAGE30US")

# choose a better column name than FRED id
rate <- rate |>
  rename(Mortgage = MORTGAGE30US)

# look at data
tail(rate, 12)
```

**Visualize:**

1. Make a time series plot of mortgage rates. Add meaningful axis labels and a title.

(A little bit of starter code is provided below, and the code chunk is set to not be evaluated with `eval: false` because the current code in there is not valid code and hence the document wouldn't render. Once you replace the code with valid code, set the chunk option to `eval: true`, or remove the `eval` option altogether since it's set to `true` by default.)

```{r}
#| eval: false

ggplot(rate, aes(___)) +
  geom___ +
  labs()
```


2. Make a time series plot of mortgage rates from the date you started high school to now.

(This time no starter code is provided, you're on your own!)


**Communicate:**

Prompt: What should someone know about the current mortgage rate compared to when you started high school?


## Stretch Exercise:

Create a time series plot of mortgage rates since 2000.
Add vertical lines to identify the election day of each US President and annotate the President's name.

(Insert R code cell and write appropriate code)
