---
title: "HW 4: Hotel Bookings"
author: YOUR NAME HERE
format: 
  html: 
    embed-resources: true
editor: visual
---

```{r}
#| label: load-packages
#| include: false
library(tidyverse)
library(scales)
```

## Hotel Pricing Demand

This dataset was created to understand hotel demand. The goal is to predict the price a consumer would purchase a hotel room. If the offered price is too high, the hotel will lose revenue because the room goes unused. If the offered price is too low, the consumer will pay for the room because it's less than they expected (got a good deal) but the hotel will have lost potential revenue because they could have charged a higher price and still got the booking.

Each row in the dataset represents a hotel booking.

Source: Antonio, N., de Almeida, A., and Nunes, L. 2019. "Hotel booking demand datasets." **Data in Brief**. 22, 41-49.

(Made popular by Tidy Tuesday)

# Import:

```{r}
hotel <- read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-11/hotels.csv')
```

# Tidy:

### Exercise 1.

Is the dataset tidy?

```{r}
slice_sample(hotel, n = 10)
```

*Note:* Your answer requires a response to each of the 3 tidy data rules.

### Exercise 2.

Do people travel on a whim?

Fill in the blanks for filtering for hotel bookings where the `lead_time` is less than 2 days.

**Note:** You will need to set `eval: true` when you have an answer you want to try out.

```{r}
#| eval: false

hotel |>
  filter(lead_time ____ ____)
```

### Exercise 3.

The properties are in the USA. Are **international guests** traveling on a whim?

Add to your code above by also including bookings where `country` is **NOT** USA.

**Note:** You will need to set `eval: true` when you have an answer you want to try out.

```{r}
#| eval: false

hotel |>
  filter(____)
```

### 

### Exercise 4.

Sometimes these late bookings are cancelled.

Add to your code above by also including bookings where `is_canceled` is 1.

**Note:** You will need to set `eval: true` when you have an answer you want to try out.

```{r}
#| eval: false

hotel |>
  filter(____)
```

### Exercise 5.

Family bookings are of interest.

Fill in the blanks for selecting the columns `children`, `babies`, `adults`, `lead_time,` `adr`.

**Note:** You will need to set `eval: true` when you have an answer you want to try out.

```{r}
#| eval: false

hotel |>
  select(____)
```

### Exercise 6.

How many bookings involve at least 1 child **or** baby?

In the following chunk, add to the code above by replacing

-   `[AT LEAST]` with the logical operator for "at least" (in two places)
-   `[OR]` with the logical operator for "or"

**Note:** You will need to set `eval: true` when you have an answer you want to try out.

```{r}
#| eval: false

hotel |>
  select(____) |>
  filter(children [AT LEAST] 1 [OR] babies [AT LEAST] 1)
```

### 

# Transform:

After discussions with the key decision makers, they want to investigate family travel to the resort property in low family travel months (specifically February and October) that don't cancel.

### Exercise 7.

Create a new dataset `familyresort` with

-   rows with

    -   `hotel` is `Resort Hotel` **AND**

    -   at least 1 child **or** baby **AND**

    -   `arrival_date_month` either February or October **AND**

    -   `is_canceled` is 0

-   columns `children`, `babies`, `adults`, `lead_time,` `arrival_date_month`, `adr`

**Note:** You will need to set `eval: true` when you have an answer you want to try out.

```{r}
#| eval: false

familyresort <- hotel |>
  filter(____) |>
  select(____) 

familyresort
```

# Visualize:

### Exercise 8.

How much are family resort travelers willing to pay for a stay? Create an appropriate graphic showing the distribution of `adr` (the Annual Daily Rate in \$) using the `familyresort` dataset.

```{r}
# add code here
# pay attention to correctness and code style
```

### Exercise 9.

Is there a difference in what guests are willing to pay between the two months? Using the `familyresort` dataset create an appropriate graphic showing the distribution of `adr` for different months.

```{r}
#| fig-cap: "Comparison of Adjusted Daily Rate (adr) for February and October."
# add code here
# pay attention to correctness and code style
```

# **Communicate:**

### Exercise 10.

Prompt: Are the booking rates different for February and October for families staying at the resort? Compare the center, spread, shape, and anything unusual.

## Stretch Exercise:

Create the dataset `cityhotel30` that includes not cancelled bookings for the City Hotel with lead times over 30 days. Create a graphic comparing `adr` for new or repeat guests (`is_repeated_guest`).

*Note:* Below is the full data dictionary. It is wide (there are lots of variables), and long (lots of rows), but most analyses will use a subset for analysis.

| variable                       | class     | description                                                                                                                                                                                                                                                                                                                                                                                                                            |
|:-----------------|:-----------------|:-----------------------------------|
| hotel                          | character | Hotel (H1 = Resort Hotel or H2 = City Hotel)                                                                                                                                                                                                                                                                                                                                                                                           |
| is_canceled                    | double    | Value indicating if the booking was canceled (1) or not (0)                                                                                                                                                                                                                                                                                                                                                                            |
| lead_time                      | double    | Number of days that elapsed between the entering date of the booking into the PMS and the arrival date                                                                                                                                                                                                                                                                                                                                 |
| arrival_date_year              | double    | Year of arrival date                                                                                                                                                                                                                                                                                                                                                                                                                   |
| arrival_date_month             | character | Month of arrival date                                                                                                                                                                                                                                                                                                                                                                                                                  |
| arrival_date_week_number       | double    | Week number of year for arrival date                                                                                                                                                                                                                                                                                                                                                                                                   |
| arrival_date_day_of_month      | double    | Day of arrival date                                                                                                                                                                                                                                                                                                                                                                                                                    |
| stays_in_weekend_nights        | double    | Number of weekend nights (Saturday or Sunday) the guest stayed or booked to stay at the hotel                                                                                                                                                                                                                                                                                                                                          |
| stays_in_week_nights           | double    | Number of week nights (Monday to Friday) the guest stayed or booked to stay at the hotel                                                                                                                                                                                                                                                                                                                                               |
| adults                         | double    | Number of adults                                                                                                                                                                                                                                                                                                                                                                                                                       |
| children                       | double    | Number of children                                                                                                                                                                                                                                                                                                                                                                                                                     |
| babies                         | double    | Number of babies                                                                                                                                                                                                                                                                                                                                                                                                                       |
| meal                           | character | Type of meal booked. Categories are presented in standard hospitality meal packages: <br> Undefined/SC -- no meal package;<br>BB -- Bed & Breakfast; <br> HB -- Half board (breakfast and one other meal -- usually dinner); <br> FB -- Full board (breakfast, lunch and dinner)                                                                                                                                                       |
| country                        | character | Country of origin. Categories are represented in the ISO 3155--3:2013 format                                                                                                                                                                                                                                                                                                                                                           |
| market_segment                 | character | Market segment designation. In categories, the term "TA" means "Travel Agents" and "TO" means "Tour Operators"                                                                                                                                                                                                                                                                                                                         |
| distribution_channel           | character | Booking distribution channel. The term "TA" means "Travel Agents" and "TO" means "Tour Operators"                                                                                                                                                                                                                                                                                                                                      |
| is_repeated_guest              | double    | Value indicating if the booking name was from a repeated guest (1) or not (0)                                                                                                                                                                                                                                                                                                                                                          |
| previous_cancellations         | double    | Number of previous bookings that were cancelled by the customer prior to the current booking                                                                                                                                                                                                                                                                                                                                           |
| previous_bookings_not_canceled | double    | Number of previous bookings not cancelled by the customer prior to the current booking                                                                                                                                                                                                                                                                                                                                                 |
| reserved_room_type             | character | Code of room type reserved. Code is presented instead of designation for anonymity reasons                                                                                                                                                                                                                                                                                                                                             |
| assigned_room_type             | character | Code for the type of room assigned to the booking. Sometimes the assigned room type differs from the reserved room type due to hotel operation reasons (e.g. overbooking) or by customer request. Code is presented instead of designation for anonymity reasons                                                                                                                                                                       |
| booking_changes                | double    | Number of changes/amendments made to the booking from the moment the booking was entered on the PMS until the moment of check-in or cancellation                                                                                                                                                                                                                                                                                       |
| deposit_type                   | character | Indication on if the customer made a deposit to guarantee the booking. This variable can assume three categories:<br>No Deposit -- no deposit was made;<br>Non Refund -- a deposit was made in the value of the total stay cost;<br>Refundable -- a deposit was made with a value under the total cost of stay.                                                                                                                        |
| agent                          | character | ID of the travel agency that made the booking                                                                                                                                                                                                                                                                                                                                                                                          |
| company                        | character | ID of the company/entity that made the booking or responsible for paying the booking. ID is presented instead of designation for anonymity reasons                                                                                                                                                                                                                                                                                     |
| days_in_waiting_list           | double    | Number of days the booking was in the waiting list before it was confirmed to the customer                                                                                                                                                                                                                                                                                                                                             |
| customer_type                  | character | Type of booking, assuming one of four categories:<br>Contract - when the booking has an allotment or other type of contract associated to it;<br>Group -- when the booking is associated to a group;<br>Transient -- when the booking is not part of a group or contract, and is not associated to other transient booking;<br>Transient-party -- when the booking is transient, but is associated to at least other transient booking |
| adr                            | double    | Average Daily Rate as defined by dividing the sum of all lodging transactions by the total number of staying nights                                                                                                                                                                                                                                                                                                                    |
| required_car_parking_spaces    | double    | Number of car parking spaces required by the customer                                                                                                                                                                                                                                                                                                                                                                                  |
| total_of_special_requests      | double    | Number of special requests made by the customer (e.g. twin bed or high floor)                                                                                                                                                                                                                                                                                                                                                          |
| reservation_status             | character | Reservation last status, assuming one of three categories:<br>Canceled -- booking was canceled by the customer;<br>Check-Out -- customer has checked in but already departed;<br>No-Show -- customer did not check-in and did inform the hotel of the reason why                                                                                                                                                                       |
| reservation_status_date        | double    | Date at which the last status was set. This variable can be used in conjunction with the ReservationStatus to understand when was the booking canceled or when did the customer checked-out of the hotel                                                                                                                                                                                                                               |
