---
title: "HW6: Forecasting Zion National Park Visitors"
author: YOUR NAME HERE
format: 
  html: 
    embed-resources: true
editor: visual
---

You are asked to predict the annual number of visitors to Zion National Park for the next five years so the Park Service can plan and prepare for the future using data from the US National Park Service.

```{r}
#| include: false
# load packages
library(tidyverse)
library(astsa)
```

## IMPORT

The US National Park Service collects an impressive amount of data on the number of visitors to our national parks. For this assignment use the National Park Service Visitor Use Statistics webpage at <https://irma.nps.gov/Stats/> for data from a selected national park.

Choose `Zion NP (ZION)` in the [**Select a Park**]{style="background-color: Linen; color: SaddleBrown;"} pull down menu. Choose *Annual Park Recreation Visitation (1919-Last Calendar Year).* Download the data as a CSV file.

Import data in an dataframe called `zion`.

Note: the first row says "Title", second row has the park, third row is blank ... the 4th row has column names, so use the `skip` argument.

**Note:** You will need to set `eval: true` when you have an answer you want to try out.

```{r}
#| eval: false
zion <- _____("_____.csv",
               # fourth row has column names
                 _____) 
```

## TIDY

**Note:** You will need to set `eval: true` when you have the dataset `zion` created in the step above.

```{r}
#| eval: false
tail(zion)
```

## TRANSFORM

Code should

-   Change the units of `RecreationVisitors` to "millions of visitors."

-   Drop the column `TotalRecreationVisitors`

**Note:** You will need to set `eval: true` when you have the dataset `zion` created in the step above.

```{r}
#| eval: false
zion <- zion |>
  mutate(RecreationVisitors = _____) |>
  _____(_____)

tail(zion)
```

## VISUALIZE

Plot the time series for Zion NP Visitors since the park's creation (1904).

**Note:** You will need to set `eval: true` when you have an answer you want to try out.

```{r}
#| eval: false
ggplot(zion, aes(x = ___, y = ___)) +
  geom_"___"() +
  geom_"___"() +
  labs(
    title = "___",
    x = "___", 
    y = "___",
    ___
    )
```

Plot the time series for Zion NP Visitors since 1950.

**Note:** You will need to set `eval: true` when you have an answer you want to try out.

```{r}
#| eval: false
ggplot(zion, aes(x = ___, y = ___)) +
  geom_"___"() +
  geom_"___"() +
  scale_y_continuous() +
  labs(
    title = "___",
    x = "___", 
    y = "___",
    ___
    )
```

## MODEL

Consider a log-transformation to satisfy the ARIMA(1,1,1) model assumptions. Plot the time series for Zion NP Visitors since 1950 with the vertical axis `log(RecreationVisitors)`

**Note:** You will need to set `eval: true` when you have the dataset `zion` created in the step above.

**Note:** You will need to set `eval: true` when you have an answer you want to try out.

```{r}
#| eval: false
ggplot(zion, aes(x = ___, y = ___)) +
  geom_"___"() +
  geom_"___"() +
  scale_y_continuous() +
  labs(
    title = "___",
    x = "___", 
    y = "___",
    ___
    )
```

Fit an ARIMA$(1, 1, 1)$ model and report the parameter estimates and standard errors.

**Note:** You will need to set `eval: true` when you have the dataset `zion` created in the step above.

```{r}
#| eval: false

# fit model and report parameter estimates and standard errors
zion1950 <- zion |>
  filter(Year > 1950) |>
  mutate(lnVisitors = log(RecreationVisitors))

zion1950_out <- sarima(zion1950$lnVisitors,1,1,1)
zion1950_out$ttable
```

Forecast the annual visitors (in millions) for the next 5 years. Report a table of predictions and 95% prediction intervals.

**Note:** You will need to set `eval: true` when you have the dataset `zion` created in the step above.

```{r}
#| eval: false

# predict log-annual visitors for next 5 years
zion1950_yhat <- sarima.for(zion1950$lnVisitors, n.ahead=5, 1,1,1)
# notice the picture
zion1950_yhat

# compute 95% prediction intervals of log-annual visitors
zion1950_yhat$L <- zion1950_yhat$pred - 2 * zion1950_yhat$se
zion1950_yhat$U <- zion1950_yhat$pred + 2 * zion1950_yhat$se

# table of un-transformed predictions for next 5 years
cbind(2023 : 2027, exp(zion1950_yhat$pred), exp(zion1950_yhat$L), exp(zion1950_yhat$U))
```

## COMMUNICATE

Where is Zion National Park? Why did you go or why do you want to?

Write a paragraph about the creation of the national park.

Did the pandemic impact visitation?

Create a graphic showing the historical values and the 5 year predictions (with uncertainty) of annual visitors (in millions).

**Note:** You will need to set `eval: true` when you have the dataset `zion` created in the step above.

```{r}
#| eval: false
plot(RecreationVisitors ~ Year, data = zion, type = "b",
     ylab = "Zion NP Annual Visitors (in millions)",
     xlim = c(1990, 2025), ylim = c(1.95, 6.7))
lines(2023 : 2027, exp(zion1950_yhat$pred), col = "darkorange2", type = "b", pch = 19)
lines(2023 : 2027, exp(zion1950_yhat$L), col = "darkorange2", lty = 2)
lines(2023 : 2027, exp(zion1950_yhat$U), col = "darkorange2", lty = 2)
```
