---
title: "Mini-Project 3: Nursery Rhyme"
author: YOUR NAME HERE
format: 
  html: 
    embed-resources: true
editor: visual
---

Select one of the following nursery rhymes; YouTube links are provided in case you don't know the lyrics.

Five little speckled frogs <https://www.youtube.com/watch?v=WSC-gHBU_d0>

Five little men in a flying saucer <https://www.youtube.com/watch?v=2E3p_51tJx0>

Five little Monkeys <https://www.youtube.com/watch?v=b0NHrFNZWh0>

There were ten in the bed <https://www.youtube.com/watch?v=2yFkQVgSBfM>

One Grey Elephant Balancing <https://www.youtube.com/watch?v=nhBqS8Nbilk>

Ten Little Fishies <https://www.youtube.com/watch?v=lf5lYxE0c5c>

## STEP ONE:

Use the `cat()` function to print out the first verse.

**Note:** You will need to set `eval: true` when you have an answer you want to try out.

```{r}
#| echo: true
#| eval: false

cat("____ 
____")
```

## STEP TWO:

Write a function that takes the number of animals as input and prints out the verse for that number of animals. Show that it works for 5 animals.

**Stretch Exercise:** Adapt your function to take a second argument that allows you to specify the type of animal

```{r}
#| echo: true
#| eval: false

____ <- function(____){

cat(
)

}

____(5)
```

## STEP THREE:

Write a loop to print the lyrics for all the verses that uses the function in STEP TWO.

```{r}
#| echo: true
#| eval: false
for
```
