---
title: "Mini-Project 1: Visualizing Star Wars Characters"
author: YOUR NAME HERE
format: 
  html: 
    embed-resources: true
editor: visual
---

```{r}
#| include: false
# load packages
library(tidyverse)
```

1.  Glimpse at the starwars data frame.

```{r}
glimpse(starwars)
```

2.  Modify the following plot to change the color of all points to `"pink"`.

```{r}
ggplot(starwars, 
       aes(x = height, y = mass, size = birth_year)) +
  geom_point(color = "royalblue")
```

3.  Add labels for title, x and y axes, and size of points. Uncomment to see the effect.

```{r}
ggplot(starwars, 
       aes(x = height, y = mass, color = gender, size = birth_year)) +
  geom_point(color = "#30509C") +
  labs(
    #title = "___",
    #x = "___", 
    #y = "___",
    #___
    )
```

4.  Pick a single numerical variable and make a histogram of it. Select a reasonable binwidth for it.

(A little bit of starter code is provided below, and the code chunk is set to not be evaluated with `eval: false` because the current code in there is not valid code and hence the document wouldn't render. Once you replace the code with valid code, set the chunk option to `eval: true`, or remove the `eval` option altogether since it's set to `true` by default.)

```{r}
#| eval: false

ggplot(starwars, aes(___)) +
  geom___
```

5.  Pick a numerical variable and a categorical variable and make a visualization (you pick the type!) to explore the relationship between the two variables. Along with your code and output, provide an interpretation of the visualization.

(This time no starter code is provided, you're on your own!)

```{r}

```

Interpretation goes here...

6.  Pick a single categorical variable from the data set and make a bar plot of its distribution.

(This time no starter code is provided, you're on your own!)

```{r}

```

7.  Pick two categorical variables and make a visualization to visualize the relationship between the two variables. Along with your code and output, provide an interpretation of the visualization.

(This time no starter code is provided, you're on your own!)

```{r}

```

Interpretation goes here...

8.  Pick two numerical variables and two categorical variables and make a visualization that incorporates all of them and provide an interpretation with your answer.

(This time no starter code is provided, you're on your own!)

```{r}

```

Interpretation goes here...
