---
title: "Visualize: Data over Time"
subtitle: BYU STAT 250
author: Dr. Scott D. Grimshaw
format: 
  revealjs:
    theme: slides.scss
    multiplex: true
    transition: fade
    slide-number: true
    chalkboard: true
    center: false
execute:
  freeze: auto
---

```{r}
#| include: false
knitr::opts_chunk$set(
  fig.width = 8,
  fig.asp = 0.818,
  fig.retina = 3,
  dpi = 300,
  out.width = "80%",
  fig.align = "center"
)
options(scipen = 100)
```


## Announcements


## Recap of previous lecture

Multivariate data analysis

-   `color` to identify another variable

-   faceting

-   table of summary statistics

## Data over time (Time Series Plot)

A single variable measured at regular time intervals

## Data: Egg Prices

google FRED Egg Prices

<https://fred.stlouisfed.org/series/APU0000708111>

Monthly Egg Price (Grade A Large)

Cost per Dozen in US City Average

## Data: Egg Prices {.smaller}

```{r}
#| echo: true
library(tidyverse)
library(scales)

eggs <- read_csv("https://fred.stlouisfed.org/graph/fredgraph.csv?id=APU0000708111")

# choose a better column name than FRED id
eggs <- eggs |>
  rename(EggPrice = APU0000708111)

# look at data
tail(eggs, 12)
```

## Time Series Plot

```{r}
#| echo: true
#| code-line-numbers: "2"
ggplot(eggs, aes(x = DATE, y = EggPrice)) +
    geom_line() 
```

## Customized Time Series Plot

::: panel-tabset
### Plot

```{r ref.label = "eggs1", echo = FALSE, warning = FALSE, out.width = "50%", fig.width = 8}
```

### Code

```{r eggs1}
#| echo: true
#| fig-show: "hide"
#| code-line-numbers: "2,8"
ggplot(eggs, aes(x = DATE, y = EggPrice)) +
    geom_line() +
    labs(
      x = NULL, y = NULL,
      title = "Monthly Egg Price",
      caption = "FRED"
    ) +
    scale_y_continuous(labels = label_dollar()) +
    theme_minimal()
```
:::

## Dates

Dates and times are as tricky in R as they were in preschool

-   7 days in a week
-   12 months in a year
-   30 days in a month (except when it's 31, or 28, or 29)
-   24 hours in a day (unless flying to NY that day)

Base R uses `POSIXlt` (seconds since January 1, 1970)

tidyverse includes `lubridate` package with objects and functions to make working with dates easier

## lubridate basics {.smaller}

date data type

```{r}
#| echo: true
today() 
```

date-time data type

```{r}
#| echo: true
now() 
```

One year ago

```{r}
#| echo: true
today() - years(1)
```

One month ago

```{r}
#| echo: true
today() - months(1)
```

Four weeks ago

```{r}
#| echo: true
today() - weeks(4)
```

## Customized Time Series Plot

::: panel-tabset
### Plot

```{r ref.label = "eggs2", echo = FALSE, warning = FALSE, out.width = "50%", fig.width = 8}
```

### Code

```{r eggs2}
#| echo: true
#| fig-show: "hide"
#| code-line-numbers: "9"
ggplot(eggs, aes(x = DATE, y = EggPrice)) +
    geom_line() +
    labs(
      x = NULL, y = NULL,
      title = "Monthly Egg Price",
      caption = "FRED"
    ) +
    scale_y_continuous(labels = label_dollar()) +
    scale_x_date(limits = c(today() - years(3), today())) +
    theme_minimal()
```
:::

## lubridate basics {.smaller}

Specifying a date

```{r}
#| echo: true
ymd("2023-11-23")
ymd("20231123")
ymd("2023/11/23")
```

Other conventions

```{r}
#| echo: true
mdy("11/23/2023")

dmy("23 Nov 2023")
```

## Customized Time Series Plot

::: panel-tabset
### Plot

```{r ref.label = "eggs3", echo = FALSE, warning = FALSE, out.width = "50%", fig.width = 8}
```

### Code

```{r eggs3}
#| echo: true
#| fig-show: "hide"
#| code-line-numbers: "10-13"
ggplot(eggs, aes(x = DATE, y = EggPrice)) +
    geom_line() +
    labs(
      x = NULL, y = NULL,
      title = "Monthly Egg Price",
      caption = "FRED"
    ) +
    scale_y_continuous(labels = label_dollar()) +
    scale_x_date(limits = c(today() - years(3), today())) +
    geom_vline(xintercept = mdy("Apr 9, 2023"), linetype="dotted") +
    annotate(
      geom = "label", x = mdy("Apr 9, 2023"), y = 1,
      label = "Easter 2023") + 
    theme_minimal()
```
:::

## Base R

```{r}
#| echo: true
ts.plot(eggs$EggPrice)
```

# Data types

## Data types in R

-   **logical**
-   **double**
-   **integer**
-   **character**
-   **data frames**, **tibbles**, and **lists** (oh my!)

------------------------------------------------------------------------

## Logical & character {.smaller}

::: columns
::: {.column width="50%"}
**logical** - boolean values `TRUE` and `FALSE`

```{r}
#| echo: true
typeof(TRUE)
```
:::

::: {.column width="50%"}
**character** - character strings

```{r}
#| echo: true
typeof("hello")
```
:::
:::

## Double & integer {.smaller}

::: columns
::: {.column width="50%"}
**double** - floating point numerical values (default numerical type)

```{r}
#| echo: true
typeof(1.335)
typeof(7)
```
:::

::: {.column width="50%"}
**integer** - integer numerical values (indicated with an `L`)

```{r}
#| echo: true
typeof(7L)
typeof(1 : 3)
```
:::
:::

## Concatenation

Vectors can be constructed using the `c()` function.

```{r}
#| echo: true
c(1, 2, 3)
c("Hello", "World!")
c(c("hi", "hello"), c("bye", "jello"))
```

## Converting between types {.smaller}

with intention...

::: columns
::: {.column width="50%"}
```{r}
#| echo: true
x <- 1 : 3
x
typeof(x)
```
:::

::: {.column width="50%"}
```{r}
#| echo: true
y <- as.character(x)
y
typeof(y)
```
:::
:::

## Converting between types {.smaller}

with intention...

::: columns
::: {.column width="50%"}
```{r}
#| echo: true
x <- c(TRUE, FALSE)
x
typeof(x)
```
:::

::: {.column width="50%"}
```{r}
#| echo: true
y <- as.numeric(x)
y
typeof(y)
```
:::
:::

## Converting between types {.smaller}

without intention...

R will happily convert between various types without complaint when different types of data are concatenated in a vector, and that's not always a great thing!

::: columns
::: {.column width="50%"}
```{r}
#| echo: true
c(1, "Hello")
c(FALSE, 3L)
```
:::

::: {.column width="50%"}
```{r}
#| echo: true
c(1.2, 3L)
c(2L, "two")
```
:::
:::

## Explicit vs. implicit coercion

Let's give formal names to what we've seen so far:

-   **Explicit coercion** is when you call a function like `as.logical()`, `as.numeric()`, `as.integer()`, `as.double()`, or `as.character()`

-   **Implicit coercion** happens when you use a vector in a specific context that expects a certain type of vector

## Data Frames and Tibbles {.smaller}

Because of the coercion to make everything the same data type, Base R has another structure that can combine objects of different data types

```{r}
#| echo: true
eggs

typeof(eggs)
```

Data Frame / Tibble is a special case (class) where every column has the same number of rows, and each column is a single data type

```{r}
#| echo: true
class(eggs)
```

*Note:* While it looks like a matrix (rows and columns), a matrix must be all numerical

## Lecture Review {.smaller}

Visualizing data over time

Dates

Data Types

\

-   Appendix:

    -   Other Material in R4DS (2e) Chapters 2, 10, 12

    -   Special Data Types

## R4DS (2e) Chapters 2, 10, 12

-   What do you think is interesting? How would you use it?
-   Sec 10.7 Coordinate Systems

## Maps

```{r}
#| echo: true
#| code-line-numbers: "1,3"
library(maps)

USstates <- map_data("state")

ggplot(USstates, aes(x = long, y = lat, group = group)) +
  geom_polygon(color = "black", fill = "lightblue")
```

## Maps

```{r}
#| echo: true
#| code-line-numbers: "5"
USstates <- map_data("state")

ggplot(USstates, aes(x = long, y = lat, group = group)) +
  geom_polygon(color = "black", fill = "lightblue") + 
  coord_quickmap() +
  theme_void()
```

# Special values

## Special values

-   `NA`: Not available
-   `NaN`: Not a number
-   `Inf`: Positive infinity
-   `-Inf`: Negative infinity

 

::: columns
::: {.column width="50%"}
```{r}
#| echo: true
pi / 0
0 / 0
```
:::

::: {.column width="50%"}
```{r}
#| echo: true
1/0 - 1/0
1/0 + 1/0
```
:::
:::

## `NA`s are wild!

```{r}
#| echo: true
x <- c(1, 2, 3, 4, NA)
```

```{r}
#| echo: true
mean(x)
mean(x, na.rm = TRUE)
summary(x)
```

## `NA`s are logical

R uses `NA` to represent missing values in its data structures.

```{r}
#| echo: true
typeof(NA)
```

## Mental model for `NA`s

-   Unlike `NaN`, `NA`s are genuinely unknown values
-   But that doesn't mean they can't function in a logical way
-   Let's think about why `NA`s are logical...

::: appex
Why do the following give different answers?
:::

::: columns
::: {.column width="50%"}
```{r}
#| echo: true
# TRUE or NA
TRUE | NA
```
:::

::: {.column width="50%"}
```{r}
#| echo: true
# FALSE or NA
FALSE | NA
```
:::
:::

## 

`NA` is unknown, so it could be `TRUE` or `FALSE`

::: columns
::: {.column width="50%"}
-   `TRUE | NA`

```{r}
#| echo: true
TRUE | TRUE  # if NA was TRUE
TRUE | FALSE # if NA was FALSE
```
:::

::: {.column width="50%"}
-   `FALSE | NA`

```{r}
#| echo: true
FALSE | TRUE  # if NA was TRUE
FALSE | FALSE # if NA was FALSE
```
:::
:::

-   Doesn't make sense for mathematical operations
-   Makes sense in the context of missing data
