---
title: "Unit 1 Drills"
author: "YOUR_NAME_HERE"
output: html_document
---

## Problem 1:

Write R code to calculate the following expression:

$$ \frac{10^3}{5} + \sqrt{49} - 8 \times 2 $$

```{r}
#| echo: true

# Write your code below to compute the result

```

## Problem 2:

Below is a table of potential variable names. Fill in the table to indicate whether each name is valid or invalid in R, and provide the reason for your answer.

| Variable Name      | Valid/Invalid | Reason |
|--------------------|---------------|--------|
| `variable1`        |               |        |
| `2variable`        |               |        |
| `my.variable`      |               |        |
| `my_variable`      |               |        |
| `if`               |               |        |
| `user-name`        |               |        |
| `.hidden_variable` |               |        |

Write your answers in the table above.

## Problem 3

Create a vector in R that contains the following sequence of numbers: 5, 10, 15, 20, 25. Use the `seq()` function.

```{r}
#| echo: true

# Create the vector using seq()

```

Now create another vector that repeats the number 3 exactly 7 times using the `rep()` function.

```{r}
#| echo: true

# Create the vector using rep()

```

## Problem 4

Given two vectors:

```{r}
#| echo: true

# Provided vectors
vec1 <- c(2, 4, 6, 8)
vec2 <- c(1, 3, 5, 7)
```

Perform the following operations in R:

- Add the two vectors. 
- Subtract vec2 from vec1. 
- Multiply the elements of the two vectors together.

```{r}
#| echo: true
# Write your code below to compute the results

```

## Problem 5

Given the vector:

``` {r}
#| echo: true

heights <- c(168, 172, 171, 169, 170)
```

Using built-in R functions

- Compute the mean height.
- Compute the standard deviation of the heights.

``` {r}
#| echo: true


```


