---
title: "Unit 1 Drills Solutions"
format: html
editor: visual
---

## Problem 1

Write R code to calculate the following expression:

$$ \frac{10^3}{5} + \sqrt{49} - 8 \times 2 $$

```{r}
#| echo: true

# Write your code below to compute the result
10^3/5+sqrt(49)-8*2
```

## Problem 2

Below is a table of potential variable names. Fill in the table to indicate whether each name is valid or invalid in R, and provide the reason for your answer.

| Variable Name      | Valid/Invalid | Reason                       |
|--------------------|---------------|------------------------------|
| `variable1`        | Valid         | nothing wrong                |
| `2variable`        | Invalid       | starts with number           |
| `my.variable`      | Valid         | nothing wring                |
| `my_variable`      | Valid         | nothing wrong                |
| `if`               | Invalid       | ppecial character in R       |
| `user-name`        | Invalid       | "-" is not a valid character |
| `.hidden_variable` | Invalid       | starts with "."              |

Write your answers in the table above.

## Problem 3

Create a vector in R that contains the following sequence of numbers: 5, 10, 15, 20, 25. Use the `seq()` function.

```{r}
#| echo: true

# Create the vector using seq()
seq(start = 5,end = 25,by = 5)
```

Now create another vector that repeats the number 3 exactly 7 times using the `rep()` function.

```{r}
#| echo: true

# Create the vector using rep()
rep(3,7)
```

## Problem 4

Given two vectors:

```{r}
#| echo: true

# Provided vectors
vec1 <- c(2, 4, 6, 8)
vec2 <- c(1, 3, 5, 7)
```

Perform the following operations in R:

-   Add the two vectors.
-   Subtract vec2 from vec1.
-   Multiply the elements of the two vectors together.

```{r}
#| echo: true
# Write your code below to compute the results
# add the vectors 
print("The sum of the vectors is ")
vec1 + vec2

# Subtract the vectors
print("The difference of the vectors is")
vec2 - vec1

# Multiply the vectors
print("The product of the vectors is ")
vec1*vec2
```

## Problem 5

Given the vector:

```{r}
#| echo: true

heights <- c(168, 172, 171, 169, 170)
```

Using built-in R functions

-   Compute the mean height.
-   Compute the standard deviation of the heights.

```{r}
#| echo: true

print(paste("The mean of heights is ",mean(heights)))
print(paste("The standard deviation of heights is ",sd(heights)))
```
