---
title: "Assignment: Hypothesis Testing in R"
author: "YOUR_NAME_HERE"
output: html_document
---

## Problem 1: Student Sleep Analysis

The average time a student sleeps per night is summarized by the following data points:

6.5, 7.5, 9.0, 5.0, 6.5, 5.0, 6.0, 8.0, 4.5, 8.5, 6.5, 6.0

We will test the hypothesis:

- Null hypothesis $H_0$: Students get an average of 8 hours of sleep per night ($\mu = 8$).
- Alternative hypothesis $H_a$: Students get less than 8 hours of sleep per night ($\mu < 8$).

Significance level: $\alpha = 0.05$

### Step-by-Step Solution, Done together in class

```{r}
#| echo: true
# Step 1: Store the data in a vector
# Replace this comment with the code to create the vector


# Step 2: Calculate the sample mean and sample standard deviation
# Add code here to calculate mean and standard deviation


# Step 3: Define the hypothesized mean and sample size
# Replace this comment with code to define the null hypothesis mean and sample size


# Step 4: Calculate the test statistic
# Add code to calculate the t-statistic



# Step 5: Compute the p-value for a one-tailed test



# Step 6: Make a conclusion based on p-value and significance level




```

## Problem 2

The average daily caffeine consumption (in mg) of a group of college students is recorded as follows:

400, 250, 0, 300, 0, 600, 250, 200, 0, 275

Perform a hypothesis test to determine if the students consume more than 260 mg of caffeine on average.

- State the null and alternative hypotheses.
- Use $\alpha = 0.05$ as the significance level.
- Compute the test statistic, p-value, and make a conclusion.

Use the provided data and fill in the steps based on what you've learned from the first scenario.

```{r}
#| echo: true

```

## Problem 3

A two sample t-test is used for a hypothesis test that the mean of two groups are equal. If the two groups are equal size, the test statistic can be found using the formula:
   $$
   t = \frac{\bar{x}_A - \bar{x}_B}{s_p \sqrt{\frac{2}{n}}}
   $$
   where $s_p$ is the pooled standard deviation:
   $$
   s_p = \sqrt{\frac{(n - 1)s_A^2 + (n - 1)s_B^2}{2n - 2}}
   $$
$\bar{x}_A$ and $\bar{x}_B$ are the sample means for groups A and B respectively and $s_A$ and $s_B$ are sample standard deviations of the respective groups. 

The table below provides the test scores of students in two groups:

| Group A | Group B |
|---------|---------|
| 85      | 88      |
| 87      | 84      |
| 83      | 89      |
| 86      | 87      |
| 84      | 85      |

For a two sample t-test, the null and alternative hypotheses are

- Null hypothesis $H_0$: The mean scores of the two groups are equal ($\mu_A = \mu_B$).
- Alternative hypothesis $H_a$: The mean scores of the two groups are not equal ($\mu_A \neq \mu_B$).

We will use significance level: $\alpha = 0.05$

Perform a two-sample balanced t-test to compare the means of the two groups. 
4. Compute the p-value for a two-tailed test.
5. Compare the p-value with $\alpha$ and make a conclusion.

```{r}
#| echo: true

```
