---
title: "Assignment: Hypothesis Testing in R"
author: "YOUR_NAME_HERE"
output: visual
format: html
editor: 
  markdown: 
    wrap: 72
---

## Problem 1

The average time a student sleeps per night is summarized by the
following data points:

6.5, 7.5, 9.0, 5.0, 6.5, 5.0, 6.0, 8.0, 4.5, 8.5, 6.5, 6.0

We will test the hypothesis:

-   Null hypothesis $H_0$: Students get an average of 8 hours of sleep
    per night ($\mu = 8$).
-   Alternative hypothesis $H_a$: Students get less than 8 hours of
    sleep per night ($\mu < 8$).

Significance level: $\alpha = 0.05$

(done together in class)

```{r}
#| echo: true
# Step 1: Store the data in a vector
sleep_hours <- c(6.5, 7.5, 9.0, 5.0, 6.5, 5.0, 6.0, 8.0, 4.5, 8.5, 6.5, 6.0)


# Step 2: Calculate the sample mean and sample standard deviation
sample_mean <- mean(sleep_hours)
sample_sd <- sd(sleep_hours)


# Step 3: Define the hypothesized mean and sample size
mu_0 <- 8
n <- length(sleep_hours)


# Step 4: Calculate the test statistic
test_statistic <- (sample_mean - mu_0)/(sample_sd / sqrt(n))

# Step 5: Compute the p-value for a one-tailed test
p_value <- pt(test_statistic,n-1)

# Step 6: Make a conclusion based on p-value and significance level
print("Because the p value is less than the significance level, we reject the null hypothesis and conclude that the true number of sleep hours in less than 8.")


# Reject the null hypothesis and conclude that the true number of sleep hours is less than 8


```

## Problem 2

The average daily caffeine consumption (in mg) of a group of college
students is recorded as follows:

400, 250, 0, 300, 0, 600, 250, 200, 0, 275

Perform a hypothesis test to determine if the students consume more than
260 mg of caffeine on average.

-   State the null and alternative hypotheses.
-   Use $\alpha = 0.05$ as the significance level.
-   Compute the test statistic, p-value, and make a conclusion.

Use the provided data and fill in the steps based on what you've learned
from the first scenario.

```{r}
#| echo: true

# Step 1: Store the data in a vector
caffeine <- c(400, 250, 0, 300, 0, 600, 250, 200, 0, 275)

# Step 2: Calculate the sample mean and sample standard deviation
x_bar <- mean(caffeine)
s <- sd(caffeine)

# Step 3: Define the hypothesized mean and sample size
mu_0 = 260
n <- length(caffeine)

# Step 4: Calculate the test statistic
test_statistic <- (x_bar - mu_0)/(s/sqrt(n))

# Step 5: Compute the p-value for a one-tailed test
p_value <- 1-pt(test_statistic,n-1)

# Step 6: Make a conclusion based on p-value and significance level
sprintf("We compute a p-value of %.3f which is larger than our alpha, and so we fail to \n reject the null hypothesis and conclude that caffeine consumption \n could reasinably be %d mg",p_value,mu_0)

# Fail to reject


```

## Problem 3

A two sample t-test is used for a hypothesis test that the mean of two
groups are equal. If the two groups are equal size, the test statistic
can be found using the formula: $$
   t = \frac{\bar{x}_A - \bar{x}_B}{s_p \sqrt{\frac{2}{n}}}
   $$ where $s_p$ is the pooled standard deviation: $$
   s_p = \sqrt{\frac{(n - 1)s_A^2 + (n - 1)s_B^2}{2n - 2}}
   $$ $\bar{x}_A$ and $\bar{x}_B$ are the sample means for groups A and
B respectively and $s_A$ and $s_B$ are sample standard deviations of the
respective groups.

The table below provides the test scores of students in two groups:

| Group A | Group B |
|---------|---------|
| 85      | 88      |
| 87      | 84      |
| 83      | 89      |
| 86      | 87      |
| 84      | 85      |

For a two sample t-test, the null and alternative hypotheses are

-   Null hypothesis $H_0$: The mean scores of the two groups are equal
    ($\mu_A = \mu_B$).
-   Alternative hypothesis $H_a$: The mean scores of the two groups are
    not equal ($\mu_A \neq \mu_B$).

We will use significance level: $\alpha = 0.05$

Perform a two-sample balanced t-test to compare the means of the two
groups. 4. Compute the p-value for a two-tailed test. 5. Compare the
p-value with $\alpha$ and make a conclusion.

```{r}
#| echo: true

```

## Problem 4

Write R code to calculate the following expression:

$$ \frac{10^3}{5} + \sqrt{49} - 8 \times 2 $$

```{r}
#| echo: true

# Write your code below to compute the result

```

## Problem 5

Below is a table of potential variable names. Fill in the table to
indicate whether each name is valid or invalid in R, and provide the
reason for your answer.

| Variable Name      | Valid/Invalid | Reason |
|--------------------|---------------|--------|
| `variable1`        |               |        |
| `2variable`        |               |        |
| `my.variable`      |               |        |
| `my_variable`      |               |        |
| `if`               |               |        |
| `user-name`        |               |        |
| `.hidden_variable` |               |        |

Write your answers in the table above.

## Problem 6

Create a vector in R that contains the following sequence of numbers: 5,
10, 15, 20, 25. Use the `seq()` function.

```{r}
#| echo: true

# Create the vector using seq()

```

Now create another vector that repeats the number 3 exactly 7 times
using the `rep()` function.

```{r}
#| echo: true

# Create the vector using rep()

```

## Problem 7

Given two vectors:

```{r}
#| echo: true

# Provided vectors
vec1 <- c(2, 4, 6, 8)
vec2 <- c(1, 3, 5, 7)
```

Perform the following operations in R:

-   Add the two vectors.
-   Subtract vec2 from vec1.
-   Multiply the elements of the two vectors together.

```{r}
#| echo: true
# Write your code below to compute the results

```

## Problem 8

Given the vector:

```{r}
#| echo: true

heights <- c(168, 172, 171, 169, 170)
```

Using built-in R functions

-   Compute the mean height.
-   Compute the standard deviation of the heights.

```{r}
#| echo: true


```
