---
title: "Homework 8"
format: pdf
editor: visual
---

## Problem 1

Using the built-in `mtcars` dataset, create a scatterplot with the following specifications:

-   Plot **weight** (`wt`) on the x-axis and **miles per gallon** (`mpg`) on the y-axis.
-   Map the following aesthetics:
-   **Color**: use `cyl` (number of cylinders) as a factor.
-   **Shape**: use `am` (transmission; 0 = automatic, 1 = manual) as a factor.
-   **Size**: use horsepower (`hp`).
-   Add informative axis labels, a title, and a caption.

```{r}
#| echo: true
library(ggplot2)
# your code here
```

*In one or two sentences, briefly describe what the plot reveals about the relationship among weight, mpg, horsepower, cylinders, and transmission type.*

## Problem 2

Create two faceted versions of a scatterplot for `mpg` versus `wt` using the `mtcars` dataset:

1.  **Facet Wrap**:\
    Facet the scatterplot by the number of cylinders (`cyl`).

```{r}
#| echo: true

```

2.  **Facet Grid**:\
    Create a grid of plots with rows corresponding to the number of cylinders (`cyl`) and columns corresponding to the number of gears (`gear`).

For both plots, include appropriate axis labels and titles.

```{r}
#| echo: true

```

*In a few sentences, compare the facet_wrap and facet_grid approaches and discuss when each might be more appropriate.*

## Problem 3

Using the `mtcars` dataset, create a scatterplot matrix of the following variables:\
- Miles per gallon (`mpg`) - Displacement (`disp`) - Horsepower (`hp`) - Weight (`wt`)

Map the `cyl` variable (as a factor) to the color aesthetic using the **GGally** package's `ggpairs()` function. Customize the color scale to use `"darkorange"`, `"purple"`, and `"cyan4"`.

```{r}
#| echo: true

# Load the GGally package for the scatterplot matrix
library(GGally)

```

*Briefly summarize one interesting relationship you observe among these variables.*

## Problem 4

Using the **table1** package, produce a summary statistics table for the `mtcars` dataset that shows the variables:

-   Miles per gallon (`mpg`)
-   Weight (`wt`)
-   Horsepower (`hp`)
-   Displacement (`disp`)

Display these statistics separately by the number of cylinders (`cyl`), and include an overall column labeled `"Total"`. Add a caption and a footnote indicating the data source.

```{r}
#| echo: true

# Load the table1 package
library(table1)

```

*In one sentence, explain how the summary statistics table helps you understand differences among cars with different numbers of cylinders.*

## Problem 5

The built-in `Orange` dataset records the growth (circumference) of orange trees measured at different ages. Create a line plot with the following specifications:

-   Plot **age** on the x-axis and **circumference** on the y-axis.
-   Draw one line per tree, using a different color for each tree.
-   Add appropriate axis labels, a title, and a caption.

```{r}
#| echo: true

```

*In one sentence, describe what the plot suggests about the growth patterns of the trees.*

## Problem 6

Instead of drawing one line per tree, use a facet grid to plot the data and add a smoothed line, one tree per facet.

```{r}
#| echo: true

```
