---
title: "Random Sampling and Data Structures: Practice Problems"
author: "YOUR_NAME_HERE"
output: html_document
---

## Problem 1

Use the sample() function to complete the following tasks:

-   Randomly reorder the names in the vector names below.
-   Select 4 names without replacement.
-   Select 5 names with replacement.

```{r}
#| echo: true
names <- c("Alice", "Bob", "Charlie", "Diana", "Ethan", "Fiona", "George", "Helen")
# Write your code here

```

## Problem 2

Perform the following tasks:

-   Create a 3x4 matrix where the elements are the numbers 1 through 12, filled row-wise.
-   Extract the element in the 2nd row, 3rd column.
-   Extract the 2nd column.

```{r}
#| echo: true
# Write your code here

```

## Problem 3

Given the matrix below, use logical masks to extract:

-   All rows where the value in the first column is greater than 2.
-   The elements in the second column that are even numbers.

```{r}
#| echo: true
matrix_example <- matrix(1:12, nrow = 4, byrow = TRUE)
matrix_example
# Write your code here
```

## Problem 4

Given the list below, complete the following tasks:

-   Extract the second element as a list.
-   Extract the contents of the second element.
-   Extract the third element and subset its first two rows.

```{r}
#| echo: true
example_list <- list(
  vector = c(10, 20, 30),
  matrix = matrix(1:9, nrow = 3),
  data_frame = data.frame(ID = 1:3, Value = c("A", "B", "C"))
)
# Write your code here
```

## Problem 5:

Given the vector below:

Create a logical mask to identify elements greater than 15. Use the mask to subset the vector. Count how many elements are greater than 15.

```{r}
#| echo: true
vec <- c(10, 25, 5, 18, 30, 12, 20)
# Write your code here
```

## Problem 6

Create a logical mask to check which rows have the Name column equal to "Alice" or "Charlie". Use the mask to subset the data frame and show only those rows.

```{r}
#| echo: true
df_example <- data.frame(
  ID = 1:5,
  Name = c("Alice", "Bob", "Charlie", "Diana", "Ethan"),
  Age = c(25, 30, 35, 40, 28)
)
# Write your code here
```
