---
title: "Homework 5: Simulation for statistical research"
format: pdf
editor: visual
---

## Problem 1

Consider the estimation of the parameter $\mu$ in a normal distribution. Use a simulation study to answer the question, which estimator for $\mu$ is "better"?

-   Mean of a sample
-   Median of a sample
-   The average of the largest and smallest value

Generate data sets of size 100, with a mean of 0, and a variance of 1 to answer this question.

```{r}
#| echo: true


```

## Problem 2

What happens in the simulation from problem 1 when you change

1.  The mean of the data generating distribution
2.  The variance of the data generating distribution
3.  The number of data points in each data set.

Write code to explore this and make a statement as to if your answer changes as to which estimator is best.

```{r}
#| echo: true


```

## Problem 3

Conduct a simulation study to assess how often a two-sample t-test correctly detects when two groups have the same or different means. Follow these two scenarios:

1.  **Equal Means:** Generate 30 samples for Group A and Group B from a normal distribution with mean 12 and standard deviation 3.
2.  **Different Means:** Generate 30 samples for Group A from a normal distribution with mean 12 and standard deviation 3, while Group B has mean 10 and standard deviation 3.

For each scenario, perform a simulation study and determine how often the t-test correctly identifies: - The means are the same in the first scenario. - The means are different in the second scenario.

```{r}
#| echo: true

## Scenario 1
# Simulated data

alpha <- .05
sim_scenario <- function(nA = 30, nB = 30,muA = 12,muB = 12,sA = 3, sB = 3){
  groupA <- rnorm(nA, mean = muA, sd = sA)
  groupB <- rnorm(nB, mean = muB, sd = sB)

# Perform t-test
  results <- t.test(groupA, groupB, var.equal = TRUE)
  results$p.value > alpha
}

B <- 10000
results1 <- replicate(B,sim_scenario())
print(paste("The proportion of correct conclusion in scenario 1 is",mean(results1)))
results2 <- replicate(B,sim_scenario(muB = 10))
print(paste("The proportion of correct conclusion in scenario 2 is",1-mean(results2)))

```

## Problem 4

Repeat the same simulation study as in Problem 3, but this time modify Group B to have a standard deviation of 1 in both scenarios.

Compare the results with those from Problem 3 and discuss the impact of unequal variances on the accuracy of the t-test.

```{r}
#| echo: true
results1 <- replicate(B,sim_scenario(sB = 2, sA = 100))
print(paste("The proportion of correct conclusion in scenario 1 is",mean(results1)))
results2 <- replicate(B,sim_scenario(muB = 10, sB = 2, sA = 1))
print(paste("The proportion of correct conclusion in scenario 2 is",1-mean(results2)))

```

## Problem 5

Conduct a simulation study to confirm the **Central Limit Theorem (CLT)** by drawing samples from a uniform distribution and analyzing the distribution of the sample means.

### Steps:

1.  Generate multiple samples from a **Uniform(0,1)** distribution.
2.  Compute the sample mean for each iteration.
3.  Use the **Shapiro-Wilk test** to assess whether the distribution of sample means is approximately normal.

Run this simulation for different sample sizes (e.g., n = 5, 10, 30, 100) and observe how the distribution of the sample means changes.

```{r}
#| echo: true

```
