---
title: "Simulation: Statistical Methods"
subtitle: "BYU STAT 250"
author: "Dr. Robert Richardson"
format: 
  revealjs:
    theme: slides.scss
    multiplex: true
    transition: fade
    slide-number: true
    chalkboard: true
    center: false
    html-math-method: mathjax
execute:
  freeze: auto
---

```{r}
#| include: false
knitr::opts_chunk$set(
  fig.width = 8,
  fig.asp = 0.818,
  fig.retina = 3,
  dpi = 300,
  out.width = "80%",
  fig.align = "center"
)
options(scipen = 100, width = 70)
```

## Intro to Statistical Research using Simulation Studies

What statistical principles do you remember from Stat 121?

-   Central Limit Theorem
-   Law of Large Numbers
-   Confidence Intervals
-   Hypothesis Tests

Can we use hypothesis tests to confirm some of these principles?

## Intro to Statistical Research using Simulation Studies

Theoretically, 68% of data from a normal distribution is within 1 standard deviation of the mean. How can we test this using a simulation study?

## Intro to Statistical Research using Simulation Studies

Our strategy could be

1.  Simulate data from a normal distribution
2.  Compute the proportion of data within 1 standard deviation.
3.  Collect the proportion inside a vector
4.  Repeat Steps 1 through 3 many times

## 68% Simulation {.smaller}

::: panel-tabset
### n = 10

```{r}
#| echo: true
# Parameters
n <- 10   # Sample size per simulation
simulations <- 10000  # Number of simulations

# Run the simulation study
proportions <- replicate(simulations, {
  sample_data <- rnorm(n, mean = 0, sd = 1)  # Generate a normal sample
  within_1sd <- mean(sample_data > -1 & sample_data < 1)  # Proportion within 1 SD
  return(within_1sd)
})

# Summary statistics
mean_proportion <- mean(proportions)
sd_proportion <- sd(proportions)

# Print results
mean_proportion

```

### n = 100

```{r}
#| echo: true
# Parameters
n <- 100   # Sample size per simulation
simulations <- 10000  # Number of simulations

# Run the simulation study
proportions <- replicate(simulations, {
  sample_data <- rnorm(n, mean = 0, sd = 1)  # Generate a normal sample
  within_1sd <- mean(sample_data > -1 & sample_data < 1)  # Proportion within 1 SD
  return(within_1sd)
})

# Summary statistics
mean_proportion <- mean(proportions)
sd_proportion <- sd(proportions)

# Print results
mean_proportion

```

### n = 1000

```{r}
#| echo: true
# Parameters
n <- 1000   # Sample size per simulation
simulations <- 10000  # Number of simulations

# Run the simulation study
proportions <- replicate(simulations, {
  sample_data <- rnorm(n, mean = 0, sd = 1)  # Generate a normal sample
  within_1sd <- mean(sample_data > -1 & sample_data < 1)  # Proportion within 1 SD
  return(within_1sd)
})

# Summary statistics
mean_proportion <- mean(proportions)
sd_proportion <- sd(proportions)

# Print results
mean_proportion

```
:::

## 68% Simulation {.smaller}

The strategy of creating a dataset in a certain way and creating a **statistic** based on the data and checking against either a theoretical value or an alternative method. For example, do you think the 68% rule will work when using a uniform distribution instead?

## 68% Simulation {.smaller}

::: panel-tabset
### n = 10

```{r}
#| echo: true
# Parameters
n <- 10   # Sample size per simulation
simulations <- 10000  # Number of simulations

# Run the simulation study
proportions <- replicate(simulations, {
  sample_data <- rexp(n, rate = 1)  # Generate an exponential sample
  within_1sd <- mean(sample_data > 0 & sample_data < 2)  # Proportion within 1 SD
  return(within_1sd)
})

# Summary statistics
mean_proportion <- mean(proportions)
sd_proportion <- sd(proportions)

# Print results
mean_proportion

```

### n = 100

```{r}
#| echo: true
# Parameters
n <- 100   # Sample size per simulation
simulations <- 10000  # Number of simulations

# Run the simulation study
proportions <- replicate(simulations, {
  sample_data <- rexp(n, rate = 1)  # Generate an exponential sample
  within_1sd <- mean(sample_data > 0 & sample_data < 2)  # Proportion within 1 SD
  return(within_1sd)
})

# Summary statistics
mean_proportion <- mean(proportions)
sd_proportion <- sd(proportions)

# Print results
mean_proportion

```

### n = 1000

```{r}
#| echo: true
# Parameters
n <- 1000   # Sample size per simulation
simulations <- 10000  # Number of simulations

# Run the simulation study
proportions <- replicate(simulations, {
  sample_data <- rexp(n, rate = 1)  # Generate an exponential sample
  within_1sd <- mean(sample_data > 0 & sample_data < 2)  # Proportion within 1 SD
  return(within_1sd)
})

# Summary statistics
mean_proportion <- mean(proportions)
sd_proportion <- sd(proportions)

# Print results
mean_proportion

```
:::

## Research Question

Many research questions can be addressed via simulation study. For example

1.  Confirming a theoretical property
2.  Determining the value of a specific method (i.e. is Method A better than Method B?)
3.  Determining the effect of violating model assumptions in a specific way

## Research Question

Consider the estimation of the parameter $\mu$ in a normal distribution. Here's a question we can answer with a simulation. Which estimator for $\mu$ is "better"?

-   Mean of a sample
-   Median of a sample
-   The average of the largest and smallest value

## "Better"

What does "Better" mean?

## Simulation Set-up Decisions

The quality of your results depend on several factors

- How good is your question?
- How are you generating your data?
- What metric are you using?

## Comparing different $\mu$'s

Question: Which method of estimating $\mu$ results in an estimate that is on average closest to the truth

Data Generation: Generate $n = 100$ values from a standard normal $N(0,1)$

Metric: We will use absolute bias, which is the absolute value of the estimate versus the truth. $Bias = | \mu_{Truth} - \hat{\mu} |$




