---
title: "Drills 7"
format: pdf
editor: visual
---

```{r}
library(ggplot2)
```

## Drill 1: Changing Binwidth in a Histogram

Below is the code for a ggplot2 histogram of diamond prices using a binwidth of 500.

**Question:** Modify the code to change the binwidth to 200.

```{r}
#| echo: true
ggplot(diamonds_ideal, aes(x = price)) +
  geom_histogram() +
  labs(x = "Price (USD)", y = "Frequency", title = "Histogram of Diamond Prices")
```

## Drill 2: Modifying a Density Plot
  
Below is a density plot code for diamond prices.

**Question:** Change the fill color to `"steelblue"` and increase the opacity from 0.7.

```{r}
#| echo: true

ggplot(diamonds_ideal, aes(x = price)) +
  geom_density(adjust = 0.9) +
  labs(x = "Price (USD)", y = "Density", title = "Density Plot of Diamond Prices")
```

## Drill 3: Fixing an Incorrect Scatterplot
  
Below is an incorrect Base R scatterplot code, but it throws an error
 Identify and fix the error.

```{r}
#| echo: true
#| eval: false
# After fixing code, change above to eval: true
# Incorrect code:
plot(diamonds$car, diamonds$price,
     pch = 19, col = "darkorange",
     xlab = "Carat", ylab = "Price (USD)",
     main = "Diamond Carat vs. Price")
```

## Drill 4: Enhancing a Q-Q Plot
  
  Below is the code for a Q-Q plot assessing the normality of diamond prices. Add a title "Corrected Q-Q Plot for Diamond Prices" to the plot.

```{r}
ggplot(diamonds_ideal, aes(sample = price)) +
  geom_qq(color = "darkorange", size = 2) +
  geom_qq_line(color = "royalblue")
```

## Drill 5: Modifying a Boxplot
  
Below is the code for a ggplot2 boxplot of diamond prices by clarity.
Modify the code to change the box fill color to `"green"` and remove the outlier points.

```{r}
ggplot(diamonds, aes(x = clarity, y = price)) +
  geom_boxplot(color = "black", fill = "darkorange") +
  labs(x = "Clarity", y = "Price (USD)", title = "Diamond Prices by Clarity")
```

