---
title: "Homework 7: Plotting univariate and bivariate data"
format: pdf
editor: visual
---

## Problem 1

In this problem, you will work with the `diamonds` dataset from the **ggplot2** package. First, subset the data to include only diamonds with an **Ideal** cut using base R logical subsetting. Then complete the following tasks:

1.  **ggplot2 Histogram**:
    -   Create a histogram of the variable `price` using a binwidth of 500.
    -   Customize the plot by setting black borders and a darkorange fill.
    -   Add appropriate x-axis and y-axis labels and a title.

```{r}
#| echo: true
# Load necessary library for the diamonds dataset and ggplot2 functions
library(ggplot2)

# Write code here
```

2.  **Base R Histogram**:
    -   Using the same subset of data, produce a histogram with 10 breaks.
    -   Customize it by adding an x-axis label, y-axis label, and main title.

```{r}
#| echo: true
# Write code here
```

## Problem 2

Using the same `diamonds_ideal` dataset, complete the following tasks:

1.  **Density Plot with ggplot2**:\

-   Create a density plot for `price` with an adjust value of 0.9.
-   Customize the plot with a darkorange fill at 50% opacity and limit the x-axis to the range 300 to 8000.

```{r}
#| echo: true

```

2.  **Overlay Normal Density**:\

-   On the same plot, overlay a Normal probability density function using `stat_function()`.
-   Use the sample mean and standard deviation for the Normal curve and set its color to royalblue.

```{r}
#| echo: true

```

3.  **Q-Q Plot**:\

-   Produce a Q-Q plot using ggplot2 (with `geom_qq()` and `geom_qq_line()`) to assess the normality of `price`.

```{r}
#| echo: true

```

4.  **Normality Tests**:\

-   Perform a Shapiro-Wilk test and a Kolmogorov-Smirnov test for normality on `price` and print the results.

```{r}
#| echo: true

```

## Problem 3

Investigate the relationship between `carat` and `price` using the full `diamonds` dataset.

1.  **ggplot2 Scatterplot**:\

-   Create a scatterplot with `carat` on the x-axis and `price` on the y-axis.
-   Customize the points (set size and color) and add a smoothing curve using `geom_smooth()` with a span of 1.0.

```{r}
#| echo: true

```

2.  **Base R Scatterplot**:\

-   Create a scatterplot of the same variables using Base R with a customized point shape and labels.

```{r}
#| echo: true

```

## Problem 4

Using the full `diamonds` dataset, complete the following tasks:

1.  **Overlaid Density Plot**:\

-   Create an overlaid density plot of `price` by `cut` using ggplot2.
-   Use custom fill colors for each level of `cut` (e.g., darkorange, darkorchid, cyan4, gold, steelblue) and adjust the transparency so the curves can be compared.

```{r}
#| echo: true

```

2.  **Bar Plots for Categorical Data**:\

-   Create a bar plot of the variable `color` to show the frequency of diamonds for each color using ggplot2.
-   Then modify the bar plot to display the relative frequencies (proportions).

```{r}
#| echo: true

```

3.  **Pareto Chart**:\

-   Produce a Pareto chart for `color` by sorting the bars in descending order of frequency.

```{r}
#| echo: true

```

## Problem 5

Using the built-in `CO2` dataset, complete the following tasks:

- Visualize the uptake variable with a fitted density overlaid. Use good labels. 

```{r}
#| echo: true

```

- Visualize the relationship between concentration and uptake. Use good labels. 

```{r}
#| echo: true

```



