---
title: "Homework 9"
format: pdf
editor: visual
---

## **Problem 1**

Read the data file <https://drbob-richardson.github.io/Stat250_W2025/data/world_happiness_wide.csv>and inspect the data.

```{r}

```

1.  Explain what makes a dataset "tidy" and why tidy data is useful for data analysis.
2.  Is the World Happiness Report dataset in tidy format?

------------------------------------------------------------------------

## **Problem 2**

1\. Convert this wide-format dataset into **tidy format** where all years are in a single column named `Year` with corresponding happiness scores in a column called `Happiness_Score`.

```{r}

```

2\. Create a plot showing happiness over time where each country has a different color.

```{r}

```

3\. Why was pivoting important in order to make this plot?

------------------------------------------------------------------------

## **Problem 3**

Perform the following operations in a single statement chained together with pipes

1.  a\. Filter the dataset to include only the most recent available year.

    b\. Further filter the dataset to include only countries where the `Happiness_Score` is greater than **6.5**.

    c\. Select only the `Country` and `Happiness_Score` variables.

```{r}

```

------------------------------------------------------------------------

## **Problem 4**

1.  Write code that groups by `Country` and prints the average `Happiness_Score` by country.

```{r}

```

2.  Write code that groups by `Year` and prints the maximum `Happiness_Score` in each year

```{r}

```

3.  Create and print a dataframe that includes one row for each `Year` along with the `Country` that has the highest `Happiness_Score` in that year.

```{r}

```

## **Problem 5**

Read in the 2019 happiness score data from <https://drbob-richardson.github.io/Stat250_W2025/data/2019.csv>

1.  Create a new column `GDP_scaled` that standardizes the `GDP_per_Capita` variable.
2.  Using `mutate()`, create a new variable called `Happiness_Index`, which is the sum of `Score`, `Social support`, `Freedom to make life choices`, and `Generosity`, minus `Perceptions of corruption`.
3.  Use `transmute()` to create a new dataset that contains only `Country`, `Year`, and the newly created `Happiness_Index`.

## **Problem 6**

1.  Find the **top 50 happiest** and **bottom 50 least happy** countries in the dataset. Create a new variable that identifies the top 50, bottom 50, or neither.

```{r}

```

2.  Create a new data set, called `Grouped_Vars` that includes the averages for each of the continuous variables in the data set grouped by if they are in the Top 50, Bottom 50, or Neither.

```{r}

```

3.  What characteristics do the happiest countries share? What about the least happy ones?
4.  Suppose a government wants to improve happiness in a country. Based on this dataset, which factors would you recommend focusing on, and why?

------------------------------------------------------------------------

## **Problem 7**

Read the data file <https://drbob-richardson.github.io/Stat250_W2025/data/global_temperatures_wide.csv>and inspect the data. These are yearly average temperatures over time. Perform the following task without manually adding or removing variables (i.e. only use tidyverse and ggplot functions):

1.  For each country and year, compute the temperature anomaly, which is defined as the difference between the observed value and the average $$anom = x_{ij} - \bar{x}_j$$ where $x_{ij}$ is the $i$-th year in the $j$-th country and $\bar{x}_j$ is the average temperature in the $j$-th country.
2.  Include only countries where the maximum anomaly over the time period is 1 or greater
3.  Plot the anomalies over time for each of those countries, each country getting a different color line.
