---
title: "Wrangling3: Your Turn"
author: YOUR NAME HERE
format: 
  html: 
    embed-resources: true
editor: visual
---

## Analysis Objective

For Disney stock returns, provide the mean and conditional variance forecast. (Could be used to value an option or other derivative.)

```{r}
#| echo: false
#| warning: false
# setup
library(tidyverse)
```

## Import

Read Historical Data from `finance.yahoo.com` website

```{r}
#| echo: true
DIS <- read_csv("https://query1.finance.yahoo.com/v7/finance/download/DIS?period1=-252374400&period2=9999999999&interval=1d&events=history&includeAdjustedClose=true")
```

## Tidy

```{r}
#| echo: true
tail(DIS, 12)
```

## Transform

Compute the log return (daily continuously compounded return

```{r}
#| echo: true
DIS <- DIS |>
  mutate(logreturn = log(Close / lag(Close)) )

tail(DIS)
```

## Visualize

```{r}
#| code-fold: true
#| echo: true
DIS |>
  filter(Date >= ymd("2019-1-1")) |>
  ggplot(aes(x = Date, y = logreturn)) +
      geom_line() +
      labs(
        x = NULL, y = "Log Return",
        title = "The Walt Disney Company (DIS)",
        caption = "\nfinance.yahoo.com"
      ) +
      scale_y_continuous(labels = scales::percent) +
      theme_minimal()
```

## Model

Fit a GARCH model to log returns

```{r}
#| echo: false
#| warning: false
# setup
library(rugarch)
```

```{r}
#| echo: true
spec <- ugarchspec()
fit1 <- ugarchfit(spec, DIS$logreturn[-1], out.sample = 20)
fit1
```

Data with +/- 2 SD

```{r}
#| echo: true
plot(fit1, which = 1)
```

## Communicate

```{r}
#| echo: true
forecast1 <- ugarchforecast(fit1, n.ahead = 20)
plot(forecast1, which = 1)
```

Table of forecast log returns and volatility

```{r}
#| echo: false
forecast1
```
