---
title: "Homework_10"
format: pdf
editor: visual
---

## Problem 1

Read in the dataset containing information on various cultural festivals: <https://drbob-richardson.github.io/Stat250_W2025/data/festivals.csv>

1.  Clean up the names of the columns to make them easier to work with

```{r}

```

2.  Convert the variable containing the date to a proper Date variable.

```{r}

```

3.  Create a new vector labeled Month and a new vector labeled Year that extracts the respect month and year variables from the Date variable. Print the first 10 rows.

```{r}

```

## Problem 2

Create a new column Type that categorizes festivals based on keywords in the Festival_Name:

-If Festival_Name contains "music", "jazz", or "rock", assign "Music". 

-If it contains "film", "cinema", or "movie", assign "Film". 

-Otherwise, assign "Other". 

-Convert the new Type column to a factor with levels "Music", "Film", "Other".

```{r}

```

## Problem 3

A second dataset, <https://drbob-richardson.github.io/Stat250_W2025/data/festival_attendance.csv>, contains attendance data with the columns Festival_Name, Year, and Attendance. Read in the festival_attendance.csv dataset. Join the attendance data with the festivals dataset using the keys Festival_Name and Year.

```{r}

```

## Problem 4

A third dataset, [https://drbob-richardson.github.io/Stat250_W2025/data/festival_s_weather.csv](https://drbob-richardson.github.io/Stat250_W2025/data/festival_weather.csv), contains weather information with columns Location, Year, and Avg_Temperature.

1.  Read in the weather dataset. Join this dataset with the merged festival_data using Location and Year as keys.

```{r}

```

2.  Convert Year to a numeric variable. Create a scatter plot to explore the relationship between Avg_Temperature and Attendance, with points colored by Year.

```{r}

```

3.  Convert Year to a factor variable and repeat the same plot.

```{r}

```

4.  Discuss, why is there a difference and which one would you prefer?

## Problem 5

Using the merged dataset from the previous problems:

1.  Create a new variable Attendance_Level based on the Attendance variable. Assign:

-   "High" if Attendance is in the top 25% quantile,

-   "Medium" if in the middle 50% quantile,

-   "Low" if in the bottom 25% quantile.


2.  Convert Attendance_Level to an ordered factor with levels "Low" \< "Medium" \< "High".

```{r}

```

3.  Create a summary table that shows the average Avg_Temperature for each combination of festival Type and Attendance_Level.

```{r}

```

## Problem 6

Look at the description of each festival.

1.  Count how many descriptions contains either the word `Independent` or the word `Indie`

```{r}

```

2.  Convert the descriptions to contain only lowercase variables. Count how many descriptions contains either the word `independent` or the word `indie`

```{r}

```

3.  Create a new variable that contains the number of words in each description.

```{r}

```
