---
title: "Wrangling 4 Your Turn: Hotel Pricing Demand"
author: YOUR NAME HERE
format: 
  html: 
    embed-resources: true
editor: visual
---

## Analysis Objective

Goal is to predict the price a consumer would purchase a hotel room

-   If offered price is too high, the hotel will lose revenue because the room goes unused.

-   If the offered price is too low, the consumer will pay for the room because they got a good deal but the hotel will have lost potential revenue because they could have charged a higher price and still got the booking.

## Import

From the Tidy Tuesday website

```{r}
#| warning: false

```

## Tidy

```{r}

```

## Understand

-   pay less by staying longer?

-   weekend stay costs more at City Hotel?

## Transform

Dataset is tall and wide. What are we focused on?

-   Room Cost

-   Focus on the City Hotel Property

-   Focus on the Bookings not cancelled

-   Do we have length of stay? How do we compute it?

-   Do we know which bookings include a weekend? How do we compute it?

```{r}

```

## Visualize

Pay Less by Staying Longer?

```{r}
#| code-fold: true

```

Weekend Stay Costs More?

```{r}
#| code-fold: true

```

## Understand

-   Learned A Lot. Need To Understand More!

-   Price is more complicated. What are other possible effects?

## Transform

Add some lines to the previous creation of `cityhotel`

-   Log transformation can't have `adr` zero

-   additional columns

```{r}

```

Create train and test datasets for modeling (70-30%)

45,073 rows in `cityhotel` (70% is 31,551.1)

```{r}

```

## Visualize

```{r}

```

## Model

Fit a regression model to `adr`

```{r}

```

Prediction Performance

```{r}

```

## Communicate

Table 1

```{r}

```

Table of forecast log returns and volatility

Signs on regression coefficients make sense

Prediction in the future can be trusted (no evidence of overfit)

Naive prediction uncertainty

$$
\hat{Y}\ \pm\ 2 \sqrt{1320} = \hat{Y}\ \pm\  2 \cdot 36.3 = \hat{Y}\ \pm\ 73
$$

***Note:*** use `predict` and untransform for precise prediction interval
