---
title: "Simulation Study Practce Problems"
format: pdf
editor: visual
---


## Problem 1

Write R code to create a vector of 10,000 zeros.

```{r}
# Create a vector of 10,000 zeros

```

## Problem 2: Using `replicate` to Repeat a Function

Below is a function that generates a random number between 0 and 1. Use the `replicate` function to repeat this function **10,000 times** and store the results in a vector.

```{r}
# Define a function that generates a random number
random_function <- function() {
  runif(1)
}


```

## Problem 3: Identifying Poisson Calculations

Below are two R code snippets. One calculates the **expected value** of a Poisson(4) random variable, while the other finds the probability that a Poisson(4) random variable is greater than 6. Identify which is which and explain why.

```{r}
A <- mean(rpois(10000, lambda = 4))


B <- mean(rpois(10000, lambda = 4) > 6)
```

## Problem 4

Write R code to generate 10,000 instances of a random variable **X**, where:

- X is the product of an **Exponential(1)** random variable and a **Bernoulli(0.5)** random variable.
- Store these values in a vector.
- Use this vector to estimate **E(X)**.


```{r}
# Set seed for reproducibility
set.seed(250)

# Define sample size
n <- 10000

# Generate the mixed random variable X
X <- rexp(n, rate = 1) * rbinom(n, size = 1, prob = 0.5)

# Estimate E(X)

```

## Problem 5

Continuing from Problem 4. Use the vector X to estimate Pr(X > 1)

```{r}
# Set seed for reproducibility
set.seed(250)

# Define sample size
n <- 10000

# Generate the mixed random variable X
X <- rexp(n, rate = 1) * rbinom(n, size = 1, prob = 0.5)

# Estimate Pr(X > 1)

```

