---
title: "Assignment: Simulations 1"
author: "YOUR_NAME_HERE"
output: 
format: pdf
---

## Problem 1

Write two functions:

1.  The function `roll2d12` rolls 2 dice with 12 sides and returns the sum
2.  The function `roll5d6` rolls 5 dice with 6 sides

(Done in class)

```{r}
#| echo: true


```

## Problem 2

Perform a simulation study to find the expected value for rolling:

1.  2 die with 12 sides
2.  5 die with 6 sides

(done together in class)

```{r}
#| echo: true




```

## Problem 3

Perform a simulation study to find the probability of rolling a 20 or more when rolling:

1.  2 die with 12 sides
2.  5 die with 6 sides

(done together in class)

```{r}
#| echo: true


```

## Problem 4

This code was part of the lecture examples:

```{r}
#| echo: true


```

Rewrite the code here using `replicate` and `sapply`

```{r}
#| echo: true

```

## Problem 5

Next semester, two sections of Stat 250 will be offered, with 60 seats available in each section.

-   There are 120 students who want to enroll in the course.
-   Each student independently has a preference - Section 1 and Section 2 with a 50% probability.
-   If more than 60 students select a section, the extra students will not get their preferred section.

Using simulation, estimate the expected number of students who will not get their preferred section.

(Done in class)

```{r}
#| echo: true

```

## Problem 6

Continuing from Problem 5, what is the probability that 5 or fewer students do not get their preference?

(Done in class)

```{r}
#| echo: true

```

## Problem 7

A student takes **two exams**, each with a different scoring distribution:

-   **Exam 1:** The score follows a **normal distribution** with
    -   Mean = **75**\
    -   Standard deviation = **10**
-   **Exam 2:** The score follows a **uniform distribution** between **60 and 100** (every value in this range is equally likely).

Use **simulation** to estimate the **expected value** of the **average score** across both exams.

That is, estimate:

$$E\left[\frac{X_1 + X_2}{2} \right]$$ where

-   $X_1 \sim N(75, 10^2)$ (Normal Distribution)
-   $X_2 \sim U(60, 100)$ (Uniform Distribution).

```{r}
#| echo: true

```

## Problem 8

Continuing from **Problem 7**.

Use **simulation** to estimate the probability that the student's **average score** is **85 or higher**, i.e.,

$$P\left(\frac{X_1 + X_2}{2} \geq 85 \right) ]$$

where

-   $X_1 \sim N(75, 10^2)$
-   $X_2 \sim U(60, 100)$ .

```{r}
#| echo: true

```
