---
title: "Stat 250 Midterm Exam: Winter 2025 - Richardson"
format: pdf
---

## Problem 1

Write an R function called `subset_between()` that takes a numeric vector and two numbers, `low` and `high`, and returns the elements of the vector that are strictly between `low` and `high` using logical subsetting.

```{r}
# Your code here
```

## Problem 2

Create a 4x4 matrix containing the numbers 1 to 16 filled by row.

```{r}
# Your code here
```

## Problem 3

Using the matrix from Problem 2, extract the submatrix consisting of rows 2 and 3 and columns 1 and 2.

```{r}
# Your code here
```

## Problem 4

Write an R function named `grade_student` that takes a single numeric score (0–100) as input and returns a letter grade according to the following rules:\
- "A" if score ≥ 90\
- "B" if 80 ≤ score \< 90\
- "C" if 70 ≤ score \< 80\
- "D" if 60 ≤ score \< 70\
- "F" if score \< 60\
Use conditional statements (if/else) in your function.

```{r}
# Your code here
```

## Problem 5

Construct a data frame named `students` with the following columns:\
- `ID`: integers from 101 to 130\
- `Score`: random integers between 50 and 100\
- `Grade`: a character vector where the grade is assigned using the function from Problem 4

```{r}
# Your code here
```

## Problem 6

Using the data frame created in Problem 4, extract the IDs where `Score` is at least 80.

```{r}
# Your code here
```

## Problem 7

Write R code to compute and print the value of the expression: $$\frac{\pi\sqrt{3}}{\log(5)}$$

```{r}
# Your code here
```

## Problem 8

Write a function that flips a fair coin 100 times. Store the outcomes in a character vector (with each element being either "Heads" or "Tails"). The function should return the proportion of "Heads" obtained.

```{r}
# Your code here
```

## Problem 9

Using the function from Problem 8, repeat the 100-coin-flip simulation 1000 times (each simulation is independent). Compute and print the simulated probability that more heads than tails are flipped.

```{r}
# Your code here
```

## Problem 10

Write a for loop that iterates over the numbers 1 through 20. For each number, if it is divisible by both 3 and 5, print "FizzBuzz"; if divisible by 3 (but not 5), print "Fizz"; if divisible by 5 (but not 3), print "Buzz"; otherwise, print the number.

```{r}
# Your code here
```
