---
title: "Unit 3 Drills"
author: "YOUR_NAME_HERE"
output: html_document
---

## Problem 1

Write a `for` loop that calculates the sum of all even numbers between 1 and 20. Print the result.

```{r}
#| echo: true

# Write your for loop here

```

## Problem 2

Write a `while` loop that starts with the value `x = 1` and repeatedly doubles `x` until it exceeds 100. Print the final value of `x`.

```{r}
#| echo: true

# Write your while loop here

```

## Problem 3

Write an R script that checks if a given number `num` is positive, negative, or zero, and prints a message for each case.

```{r}
#| echo: true

# Example number
num <- -5

# Write your if statement here

```

## Problem 4

Write an R script that checks if a given number `num` is:
- Between 10 and 20 (inclusive), or
- Greater than 50.

If either condition is true, print `"Condition met"`, otherwise print `"Condition not met"`.

```{r}
#| echo: true

# Example number
num <- 15

# Write your code here

```

## Problem 5

Write a function `multiply_by_two()` that takes a single argument, doubles its value, and returns the result. Test the function with the input `5`.

```{r}
#| echo: true

# Write your function here

```

## Problem 6

Consider the following incomplete R script. Fill in the missing code and explain the difference between the global variable `a` and the local variable `a` inside the function.

```{r}
#| echo: true

# Global variable
a <- 10

# Function
modify_a <- function() {
  a <- 20  # Local variable
  return(a)
}

# Call the function and print both values of `a`

```

## Problem 7

Write nested `for` loops to create the following pattern using asterisks:

```
*
**
***
****
*****
```

Each row of the pattern should have an increasing number of asterisks, starting with 1 in the first row and ending with 5 in the last.

```{r}
#| echo: true

# Write your nested for loops here

```
