---
title: "Homework 6: Data Collaboration and Processing"
format: pdf
editor: visual
---

## Background

You are collaborating with a colleague who sent you a data set named `collaboration_data.csv` (located [here](https://drbob-richardson.github.io/Stat250_W2025/data/collaboration_data.csv) folder). Your task is to examine the data and filter it so that only rows with a `Score` of at least 50 are kept. In this assignment you will:

1.  **Write a function** in a separate R script (`process_data.R`) that takes a data frame as input and returns a revised data frame containing only rows where the `Score` column is greater than or equal to 50.
2.  **Source** that function into this QMD file.
3.  **Load** the data set.
4.  **Process** the data using your function.
5.  **Save** the filtered data to a new file (`collaboration_data_filtered.csv`).
6.  **Save** your entire R environment so that you can send everything to your colleague.

You must turn in: 

- The PDF output of this document. 
- Your extra R script (`process_data.R`). 
- The filtered data file (`collaboration_data_filtered.csv`). 
- The saved R environment image (`collaboration_environment.RData`).

------------------------------------------------------------------------

## Problem 1: Create the Processing Function

Create a separate R script file named **process_data.R**. In that file, write a function named `process_data()` that takes one argument (a data frame) and returns a revised data frame that includes only rows where the `Score` column is at least 50.

------------------------------------------------------------------------

## Problem 2: Source the Function and Load the Data

In the code chunk below, source your `process_data.R` script.

```{r}
#| echo: true

```

## Problem 3:

In the code chunk below, load the data set `collaboration_data.csv`. It can be found at <https://drbob-richardson.github.io/Stat250_W2025/data/collaboration_data.csv>

```{r}
#| echo: true

```

## Problem 4: Process the Data

Now use the function you sourced to process the data by filtering for rows with a `Score` of at least 50. Display the first 10 observations.

```{r}
#| echo: true

```

------------------------------------------------------------------------

## Problem 5: Save the Filtered Data

Save the filtered data set to a new CSV file named `collaboration_data_filtered.csv`.

```{r}
#| echo: true


```

------------------------------------------------------------------------

## Problem 6: Save the R Environment

Finally, save your entire R environment to a file named `collaboration_environment.RData` so you can send your colleague everything you have.

```{r}
#| echo: true

```
